R Notebook

### 2016-07-25 finalwriteup/draft.rmd
started kerrie table

Ravi Kumar and Ravi [@Kumar2007] performed a detailed review of statistical and machine learning techniques that were applied over 37 years in the context of bankruptcy prediction in banks. The most widely used model was Neural Networks, however Logistic and Linear Regression, decision trees, SVM's, discriminant analysis (DA), and statistical clustering techniques (such as K Nearest Neighbour) were also popular. It was found that DA and Linear Regression techniques were not preferred due to their low accuracy. The overall assessment was that SVM's outperformed Neural Networks (back propagation Neural Networks were used most extensively), which sometimes outperformed decision trees, and the rest of the methods were generally inferior. Although SVM's performed the best, as discussed, they are often complex and slow, requiring a great deal of memory [@Kumar2007]. In further support of SVM predictive performance, a study by Davenport and Harris [@Davenport2007] concluded that many statisticians experienced in predictive machine learning algorithms agree that SVM's yield the highest predictive accuracy compared to other machine learning algorithms.

@@

Case study projects with extreme 'return per dollar' values were likely the result of special scenarios which the model is not intended to predict. For example, many projects appeared wildly profitable but on closer inspection these were very small jobs. Outliers were initially defined as 'return per dollar' values sitting outside 1.5 x the interquartile range (IQR) from the upper and lower quartiles. This rule was initially practised by Tukey, the inventor of the box plot [@Tukey1977]. All outlier projects determined by this method were reviewed with an employee from the case study company. It was concluded that outliers that were small jobs most likely required so little time that busy employees did not bother writing down their hours for that job. Other bigger outlier projects had unrealistically low hours logged against them which could have been the result of data entry errors or errors from when the data was transferred across databases. Based on the employee's domain knowledge and assessment of the outlier cases, reasonable cut-off values for the range of return per dollar values were determined. In summary, projects with 'return per dollar' values greater than 3 were removed and one project with a very low 'return per dollar' was removed because it contained incorrect invoiced amounts.

26 of the 36 outlier cases with a high 'return per dollar' were small projects worth less than \$4,000. This may indicate small projects are not suitable to predict due to timesheet entry habits on these projects, however after outlier deletion, 1,159 projects below \$4,000 remained in the dataset. The prediction model would therefore still be applicable to small projects under \$4,000.

>>>>

@Kumar2007 performed a detailed review of statistical and machine learning techniques that were applied over 37 years in the context of bankruptcy prediction in banks. The most widely used model was Neural Networks, however Logistic and Linear Regression, decision trees, SVM's, discriminant analysis (DA), and statistical clustering techniques (such as K Nearest Neighbour) were also popular. It was found that DA and Linear Regression techniques were not preferred due to their low accuracy. The overall assessment was that SVM's outperformed Neural Networks (back propagation Neural Networks were used most extensively), which sometimes outperformed decision trees, and the rest of the methods were generally inferior. Although SVM's performed the best, as discussed, they are often complex and slow, requiring a great deal of memory [@Kumar2007]. In further support of SVM predictive performance, a study by Davenport and Harris [@Davenport2007] concluded that many statisticians experienced in predictive machine learning algorithms agree that SVM's yield the highest predictive accuracy compared to other machine learning algorithms.

@@

Case study projects with extreme 'return per dollar' values could be the result of mistakes or special scenarios which the model is not intended to predict. Several methods could have been used for outlier detection including the distribution of the response variable, an analysis of high leverage cases, and cluster based methods. Guidelines for the first method were intially defined by Tukey, the inventor of the box plot [@Tukey1977]. He recommended investigating cases with response variable values sitting outside 1.5 x the interquartile range (IQR) from the upper and lower quartiles.

Alternatively, defining outliers in terms of high leverage requires solving for statistical measures such as Cook's distance or DFFITS for each case. Cook's Distance is the scaled change of all predicted values resulting from the deletion of an observation using linear or logistic regression [@Nurunnabi2009]. In contrast, DFFITS measures the change in the predicted value of a single case when the case is left out of the regression model [@Nurunnabi2009]. Cluster based outlier detection involves clustering your data via unsupervised learning, and inspecting the cases with the highest distance to the centroid of their cluster [@Schubert2014]. For this project, Tukey's method was initially chosen so that the case study company would have a simple intuitive method to follow for excluding future outlier data points. A simple method was also appealing because the outliers would be reviewed with an employee so that their domain knowledge could interpret causes for outliers and refine the rules.

Upon review of the outliers with an employee, many projects which appeared wildly profitable, were actually small jobs and high profitability ratios only indicated extra profits of a few hundred dollars. It was concluded that small job outliers most likely required so little time that busy employees did not bother writing down their hours for that job. Other bigger outlier projects had unrealistically low hours logged against them which could have been the result of data entry errors or errors from when the data was transferred across databases about 4 years prior to data extraction. Based on the employee's domain knowledge and assessment of the outlier cases, reasonable cut-off values for the range of return per dollar values were revised.

In summary, projects with 'return per dollar' values greater than 3 were removed and one project with a very low 'return per dollar' was removed because it contained incorrect invoiced amounts. 26 of the 36 outlier cases with a high 'return per dollar' were small projects worth less than \$4,000. This may indicate small projects are not suitable to predict due to timesheet entry habits on these projects, however after outlier deletion, 1,159 projects below \$4,000 remained in the dataset. The prediction model would therefore still be applicable to small projects under \$4,000.

### 2016-07-25 finalwriteup/trial_code_eq.rmd
started kerrie table

where $\boldsymbol{w}$ and $\boldsymbol{x}$ are vectors with lengths corresponding to the number of explanatory variables. For two variables, $\boldsymbol{w}$ would be $w\begin{pmatrix}\beta_1 \\ \beta_2\end{pmatrix}$ and $\boldsymbol{w}$ would be $x\begin{pmatrix}X_1 \\ X_2\end{pmatrix}$. Expanding the dot product of these two vectors gives the equation for a line:

>>>>

where $\boldsymbol{w}$ and $\boldsymbol{x}$ are vectors with lengths corresponding to the number of explanatory variables. For two variables, $\boldsymbol{w}$ would be $\begin{pmatrix}\beta_1 \\ \beta_2\end{pmatrix}$ and $\boldsymbol{w}$ would be $\begin{pmatrix}X_1 \\ X_2\end{pmatrix}$. Expanding the dot product of these two vectors gives the equation for a line:
